---
title: "PA1.Template"
author: "Dena Memari"
date: "Sunday, July 19, 2015"
output: html_document
---
## Loading and preprocessing the data
###Active Echo and read data in R: 

```{r setoption,echo = TRUE}
setwd("C:/Users/Dena/New R")
activity <- read.csv("activity.csv", header = T, sep = ",")
```
## What is mean total number of steps taken per day?
###Mean total number of steps taken per day:
```{r}
Sum <- tapply(activity$steps, activity$date, sum, na.rm=T)
```
###Histogram Total steps taken per day:
```{r}
hist(Sum, xlab = "sum of steps per day", main = "Histogram of Steps (per day)")
```
### calculate total Mean per day:
```{r}
Mean_Total <- round(mean(Sum))
print(c("Mean is",Mean_Total))
```
### Calculate total Median per day: 
```{r}
Median_Total <- round(median(Sum))
print (c("Median is", Median_Total))
````
## What is the average daily activity pattern?
###Histogram ( include Total Mean and Total Median):
### The average daily activity pattern and Activity Pattern Diagram : 
```{r}
Activity_Pattern <- tapply( activity$steps, activity$interval, mean, na.rm=T)
plot(Activity_Pattern ~ unique( activity$interval), type="l", xlab = "five Minutes Interval")
```
### maximum number of steps in 5-Mins interval:
```{r}
Activity_Pattern[which.max(Activity_Pattern)]
```
## Imputing missing values
### Calculate total missing value in whole dataset:
```{r}
table(is.na(activity) == TRUE)
```
###Strategy for filling in all of the missing values in the dataset( Create New dataSet called New_Activity):
### subsitute NA( in activity$step) with the mean (of steps) of the corresponding interval.
```{r}
New_Activity<- activity  
for (i in 1:nrow(activity)){
  if(is.na(activity$steps[i])){
    New_Activity$steps[i]<- Activity_Pattern[[as.character(activity[i, "interval"])]]
      }
   }
```
### Histogram for summary of New Dataset:
```{r}
SumNew <- tapply(New_Activity$steps, New_Activity$date, sum, na.rm=T)
hist(SumNew, xlab = "Sum of steps(per day)", main = "Histogram of steps (per day)")
```
### New Median and Mean: 
```{r}
New_Mean<- round(mean(SumNew))
New_Median <- round(median(SumNew))
print(c("The New Mean is", New_Mean))
print(c("This New Median is", New_Median))
```
## Are there differences in activity patterns between weekdays and weekends?
###different activity patterns between weekdays and weekends:
```{r}
New_Activity$weekday <- c("weekday")
New_Activity[weekdays(as.Date(New_Activity[, 2])) %in% c("Saturday", "Sunday", "samedi", "dimanche", "saturday", "sunday", "Samedi", "Dimanche"), ][4] <- c("weekend")
table(New_Activity$weekday == "weekend")
```
### Mean of activity for both weekdays and weekends:
```{r}
New_Activity$weekday <- factor(New_Activity$weekday)
New_Activity_weekend <- subset(New_Activity, New_Activity$weekday == "weekend")
New_Activity_weekday <- subset(New_Activity, New_Activity$weekday == "weekday")
Mean_New_Activity_weekday <- tapply(New_Activity_weekday$steps, New_Activity_weekday$interval, mean)
Mean_New_Activity_weekend <- tapply(New_Activity_weekend$steps, New_Activity_weekend$interval, mean)
```
### A comparison between weekdays and weekends in New dataset:
```{r}
library(lattice)
df_weekday <- NULL
df_weekend <- NULL
df_final <- NULL
Weekday <- data.frame(interval = unique(New_Activity_weekday$interval), AVG = as.numeric(Mean_New_Activity_weekday ), day = rep("weekday", length(Mean_New_Activity_weekday)))
Weekend <- data.frame(interval = unique(New_Activity_weekend$interval), AVG = as.numeric(Mean_New_Activity_weekday ), day = rep("weekend", length(Mean_New_Activity_weekend)))
Final <- rbind(Weekday, Weekend)

xyplot( AVG ~ interval | day, data = Final, layout = c(1, 2), 
        type = "l", ylab = "Number of STEPS")
```

### we can observe :
1. density of activity in weekends are higher than weekdays 
2. number of steps that are taken in weekends are lower than weekdays in general.
  
